---
title: "Food Analytics_Final Project"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r menu}
#summary(menu)
library(packrat)
```

## Including Plots

You can also embed plots, for example:

```{r, echo=FALSE}
library('ProjectTemplate')
#create.project('Food Analytics_Final Project',template="FULL",merge.strategy = "allow.non.conflict")
library(readr)
menu <- read_csv("C:/Krithika/Applied_project-R/menu.csv")               
View(menu)
head(menu)
str(menu)
library(ggplot2) # Data visualization
library(readr) # CSV file 
library(ggthemes) # extra themes for ggplot2
library(stringr) # for pattern matching
library(repr)
library(plotly)
library(tidyr)
library(dplyr)
library(tibble)
library(purrr)
```







```{r}
library(ggplot2)
ggplot(subset(menu,Category=="Breakfast"), aes(reorder(Item,Calories),Calories,fill=Calories)) + geom_bar(colour="yellow",stat = "identity")+ggtitle("Breakfast Menu Calories") + xlab("Item") + ylab("Calories")+ coord_flip()
```

```{r}
library(sqldf)
Mincal<-aggregate(menu$Calories, by=list(menu$Category), FUN=min)
colnames(Mincal)[1]<-"Category"
colnames(Mincal)[2]<-"Calories"
Mincalmenu<-sqldf("select a.Category,a.Item, a.Calories from menu a, Mincal b where a.Category==b.Category and a.Calories==b.Calories" )
## The below table gives each category minimum value
Mincalmenu
library(ggplot2)
```

##Association among variables

```{r}
library(lattice)
splom(~menu[c(4,6,11,13,15,17,19,20)],groups = NULL, data = menu,axis.line.tck = 0, axis.text.alpha = 0)
```



```{r}
menu<-na.omit(menu)
m_data<- data.frame(scale(menu[c(4:24)]))
menu1<-cbind(menu[1:3],m_data)
library(ggplot2)
qplot(Calories, Carbohydrates....Daily.Value., data=menu1, color=Category,facets=~Category, size=I(1),
      xlab="Calories", ylab="Carbohydrates....DailyValue") 

qplot(Calories, Protein, data=menu1,facets=~Category, color=Category, size=I(1),
      xlab="Calories", ylab="Protein") 

qplot(Calories, Sugars, data=menu1,facets=~Category, color=Category, size=I(1),
      xlab="Calories", ylab="Sugars")

library(corrplot)
M <- cor(m_data)
cex.before <- par("cex")
par(cex = 0.7)
corrplot(M, insig = "blank", method = "color",  tl.cex = .75)
par(cex = cex.before)

```
## % Nutrient Daily value in different category of menu
```{r}
menu <- read_csv("C:/Krithika/Applied_project-R/menu.csv")               
#View(menu)
menu %>% filter(Item %in% c("Egg McMuffin","Egg White Delight","Quarter Pounder with Cheese","Cheeseburger","Sausage, Egg & Cheese McGriddles","Premium McWrap Chicken Sweet Chili (Crispy Chicken)"))%>% 
select(Item,Cholesterol,`Sodium (% Daily Value)`,`Calcium (% Daily Value)`,Protein,`Iron (% Daily Value)`,
`Vitamin A (% Daily Value)`,`Vitamin C (% Daily Value)`,Sugars,`Saturated Fat (% Daily Value)`)%>%
gather(nut,value,2:10)%>%ggplot(aes(x="",y=value,fill=nut))+geom_bar(stat="identity",width=4)+
coord_polar(theta = "y", start=0)+facet_wrap(~Item)+theme(legend.position = "bottom",legend.text=element_text(size=5))+labs(title="Nutrient Daily value",fill="Nutrients")
```

##






###  Sodium content in each category of menu
#The American Heart Association recommends no more than 2,300 milligrams (mgs) a day and an ideal limit of no more than 1,500 mg per day for most adults. 
  
```{r}
plot_ly(x = menu$Category, y=menu$Sodium,color = menu$Category,type = "bar") %>% layout(title = "Sodium",
                                                                                                          xaxis = list(title = ""),
                                                                                                          yaxis = list(title = ""),
                                                                                                          showlegend=FALSE,
                                                                                                         autosize = T)
```
## Sugar present in each category of menu
According to the American Heart Association (AHA), the maximum amount of added sugars you should eat in a day are (7): Men: 150 calories per day (37.5 grams or 9 teaspoons). Women: 100 calories per day (25 grams or 6 teaspoons).
```{r}
plot_ly(x = menu$Category, y=menu$Sugars,color = menu$Category,type = "bar") %>% layout(title = "Sugars",
                                                                                                          xaxis = list(title = ""),
                                                                                                          yaxis = list(title = ""),
                                                                                                          showlegend=FALSE,
                                                                                                         autosize = T)
```



```{r}
rownames(menu)=as.character(menu$Item)
head(menu)
palette(rainbow(12))
stars(menu[menu$Category=="Breakfast",c(18,21:24)],key.loc=c(4,18.5),draw.segments = T)
#loading the visualization package
library(formattable)
#and also loading dplyr package, which is necessary for data transformation
library(dplyr)
```
##Average Transfat,Sugar,Sodium,Carbohydrates,Cholestrol,Calories present in menu

```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(plotly))
suppressMessages(library(RColorBrewer))
suppressMessages(library(devtools))
suppressMessages(library(yarrr))
suppressMessages(library(tidyr))
suppressMessages(library(gridExtra))
suppressMessages(library(viridis))
p <-menu %>% group_by(Category)%>%
    summarise(
      vit_a = mean(`Vitamin A (% Daily Value)`,na.rm=T),
      vit_c = mean(`Vitamin C (% Daily Value)`,na.rm=T),
      iron = mean(`Iron (% Daily Value)`,na.rm=T),
      calcium = mean(`Calcium (% Daily Value)`,na.rm=T),
      carbs = mean(`Carbohydrates (% Daily Value)`,na.rm=T),
      dietary_fiber = mean(`Dietary Fiber (% Daily Value)`,na.rm=T),
      sodium= mean(`Sodium (% Daily Value)`,na.rm=T),
      cholestrol = mean(`Cholesterol (% Daily Value)`,na.rm=T)
      
    ) %>%
    gather(type,value,2:9) %>%
    ggplot(aes(x=Category,y=value))+geom_bar(stat="identity",fill="tomato4")+
    facet_wrap(~type,ncol=4)+labs(title="Mean Percentage Daily value")+theme(axis.text.x = element_text(size=4,colour ="white"))
  p
  
```



##Average Transfat,Sugar,Sodium,Carbohydrates,Cholestrol,Calories present in menu
```{r}
 p1<-menu %>% group_by(Category)%>%
    summarise(
      sugar = mean(Sugars,na.rm=T),
     # saturatedfat = mean(`Saturated Fat`,na.rm=T)
      transfat = mean(`Trans Fat`,na.rm=T),
      sodium = mean(Sodium,na.rm=T),
      carbohydrates = mean(Carbohydrates,na.rm=T),
      cholesterol = mean(Cholesterol,na.rm=T),
      Calories = mean(Calories,na.rm=T),
      protein = mean(Protein,na.rm=T)
      
    ) %>%
    gather(type,value,2:7) %>%
    ggplot(aes(x=Category,y=(value)))+geom_bar(stat="identity",fill="tomato4")+
    facet_wrap(~type,ncol=3)+labs(title="Average Transfat,Sugar,Sodium,Carbohydrates,Cholestrol,Calories")+theme(axis.text.x = element_text(size=4))+coord_flip()
  p1
  

```

##Average Fat Percentage Daily Value in each category of menu

```{r}
p3<-menu %>% group_by(Category)%>%
    summarise(
      totalfat = mean(`Total Fat (% Daily Value)`,na.rm=T),
      saturatedfat = mean(`Saturated Fat (% Daily Value)`,na.rm=T)
      
    ) %>%
    gather(type,value,c(2,3)) %>%
    ggplot(aes(x=Category,y=(value)))+geom_bar(stat="identity",fill="hotpink")+
    facet_wrap(~type,ncol=3)+labs(title="Average Fat Percentage Daily Value")+theme(axis.text.x = element_text(size=6,colour="navyblue"),aspect.ratio = 3/4)+coord_polar()
  p3
```

## Percetage of calories from fat
```{r}
library(tidyr)
  t <- menu %>% na.omit() %>% mutate(fatpercentage = 100*(menu$'Calories from Fat'/menu$'Calories'))
 p4 <-ggplot(data=t,aes(fatpercentage))+geom_freqpoly(mapping =NULL,data=t,color="tomato4",position="identity",na.rm = FALSE)
p4

```


## Calories and 
```{r}

g1<-menu%>%ggplot(aes(x=Cholesterol,y=Calories))+geom_point(col="slateblue4")+geom_smooth(method="lm",col="slateblue4")
g2<-menu%>%ggplot(aes(x=Carbohydrates,y=Calories))+geom_point(col="tomato4")+geom_smooth(method="lm",col="tomato4")
g3<-menu%>%ggplot(aes(x=`Total Fat`,y=Calories))+geom_point(col="magenta")+geom_smooth(method="lm",col="magenta")
g3<-menu%>%ggplot(aes(x=Sugars,y=Calories))+geom_point(col="navyblue")+geom_smooth(method="lm",col="navyblue")
g4<-menu%>%ggplot(aes(x=Protein,y=Calories))+geom_point(col="firebrick4")+geom_smooth(method="lm",col="firebrick4")
g5<-menu%>%ggplot(aes(x=Sodium,y=Calories))+geom_point(col="darkorchid4")+geom_smooth(method="lm",col="darkorchid4")
g6<-menu%>%ggplot(aes(x=`Saturated Fat`,y=Calories))+geom_point(col="olivedrab4")+geom_smooth(method="lm",col="olivedrab4")
g7<-menu%>%ggplot(aes(x=`Dietary Fiber`,y=Calories))+geom_point(col="navyblue")+geom_smooth(method="lm",col="navyblue")
g8<-menu%>%ggplot(aes(x=`Trans Fat`,y=Calories))+geom_point(col="hotpink")+geom_smooth(method="lm",col="hotpink")
grid.arrange(g1,g2,g3,g4,g5,g6,g7,g8,nrow=3,ncol=4)

```
##split the dataset
```{r}
library(caTools)
set.seed(2) #to get the same split everytime
split <- sample.split(menu$Calories,SplitRatio = 0.60)
train <- subset(menu,split == "TRUE")
test <- subset(menu, split == "FALSE")
```


##Linear regression model between calories and Saturated fat.
Independent Variable - Saturated fat

Scatter plot and Conditional expectation(mean) plot
```{r}
require(dplyr)
#Scatter PLot
plot(train$`Saturated Fat`,train$Calories,main = "Scatter Plot",xlab = "Saturated Fat", ylab = "Calories")
abline(lm(train$Calories~train$`Saturated Fat`))
#Conditional Expectation Plot
dataexp <- summarise(group_by(train,`Saturated Fat`),calmean = mean(Calories))
plot(dataexp$`Saturated Fat`,dataexp$calmean,xlab = "Calories from Fat",ylab = "mean - calories",main = "Conditional 
     Expectation(mean) Plot")
```

##Linear Regression Model

```{r}
model1 <- lm(Calories~`Saturated Fat`,data = train)
summary(model1)
```


value of intercept = 128.816  value of slope =  40.328 

Both the values are significant(*** refers to high signficance) R-squared = 73% (73% of variance in calories is explained by Saturated fat)
The overall p-value is also significant.

The linear equation to predict calories : Calories =  128.816 + 40.328 *saturated fat

##Multiple regression Model
R - sq value is almost 1. These three variables almost explains 100% of variance in calories
```{r}
model <- lm(Calories~ `Saturated Fat` + Protein + Carbohydrates, data = train)
summary(model)
```
##Regression Diagnostics
```{r}
plot(model, which = 1:4)
```
```{r}
train <- train[-c(120,121,259),]
model <- lm(Calories~ `Saturated Fat` + Protein + Carbohydrates, data = train)
summary(model)
```
```{r}
predictions <- predict(model,test)
predictions
```
```{r}
plot(test$Calories,type = "l",lty = 1.8, col="navyblue")
lines(predictions,type = "l", col = "orange")
```

##Future Prediction, calories we receive=328.3043
```{r}
predict(model,data.frame('Saturated Fat'<-10,Protein = 10,Carbohydrates = 30))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
